---
title: "lab_02"
author: "derek willis"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1.**
The distribution of the expenses is that there was more in the $0-$500 range, and very few in the largest range of $100001-$500000.The larger the categories are the less expenses. I would maybe make the categories more even, like each category is a $500 difference.

```{r}
maryland_expenses <- maryland_expenses %>%
  mutate(
    
    
    divided_expenses = case_when(
        amount <=500 ~ "$0-$500",
        amount >500 & amount <5000 ~ "$500-$5000",
        amount >5000 & amount <50000 ~ "$5001-$50k",
        amount >50000 & amount <100000 ~ "$50k-$100k",
        amount >100000 & amount <500000 ~ "$100k-up",
        TRUE ~ as.character(NA)
        
        
      )
  )
```
```{r}
maryland_expenses %>%
  group_by(divided_expenses) %>%
  summarize(total = n()) %>%
  arrange(desc(total))
```

**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2.** 
The department that has the more courses with "Climate" in the title is the Department of Atmospheric and Oceanic Science (AOSC).
```{r}
umd_courses %>% 
  filter(str_detect(title, "Climate") & seats > 0) %>% 
  group_by(department) %>%
  summarize(classes = n()) %>%
  arrange(desc(classes))
```

```
**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3.** 
The most interesting department to me would be any of the humanities department such as Criminanology/Criminal Justice. It would be interesting to see how often these courses with the work "Climate" in it are taught.



**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.**

```{r}
maryland_expenses <- maryland_expenses %>% 
  mutate(expenditure_date=mdy(expenditure_date)) %>%
  mutate(month = floor_date(expenditure_date, "month")) %>%
  filter(committee_name == "Moore Wes For Maryland") %>%
  group_by(month) %>%
  summarise 

```



**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A5.** 
The question I am going to be answering is what is the most used expense method? My thought process was that I am doing to mutate all the different expense methods and then count how many times each one was used. Then I would arange that column in descending order to find out what the most used method was. Unfortunatley I am running out of time and will probably not be able to figure it out on time, but that is what I was thinking.
```{r}
maryland_expenses_with_state <- maryland_expenses %>%
  mutate(
    most_expense_method = case_when(
        str_detect(expense_method, "Check") ~ "Check",
        str_detect(expense_method, "Debit Card") ~ "Debit Card",
        str_detect(expense_method, "EFT") ~ "EFT",
        str_detect(expense_method, "Credit Card/VISA") ~ "Credit Card/VISA",

        TRUE ~ as.character(NA)
      )
  )
```
```{r}
maryland_expenses_with_state %>%
  group_by(expense_method) %>%
  summarize(total = n()) %>%
  arrange(desc(total))
```

