---
title: "Class Reference"
author: "Laura Wortman"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
wes_moore_contributions <- read_csv("https://raw.githubusercontent.com/stephenneukam/CNS_Annapolis/main/Campaign_finance/Moore_ContributionsList.csv")
```


## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2023_spring")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to count rows

You can use the 'n()' function to count rows, like so:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n())
```



```{r}
head(umd_courses)
summary(umd_courses)
colnames(umd_courses)
glimpse(umd_courses)
```
head, gives first six rows
summary, some details

filter template
```{r}
journalism_courses <- umd_courses %>% filter(department == "Journalism")

head(journalism_courses)
```
==, only journalism
!=, everything but journalism