---
title: "pre_lab_06.Rmd"
author: "derek willis"
date: "2023-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Points to hit

1.  Review of third lab questions/problems.
2.  Demonstration of combining and merging

## Chapter 23

### Task 1: Load libraries and settings

**Task** Run the following code in the gray-colored codeblock below to load the tidyverse library and turn off scientific notation.

```{r}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

### Task 2: Load data

Let's say that we have Maryland county voter registration data from four different elections in three different files. They have a similar layout and the same number of counties (plus Baltimore City).

**Task** Load the four CSV population estimate files. Do they have the same column structure? 
**Answer**
They all do have the same column structures with the same column names and datatypes.
```{r}
county_voters_2016 <- read_csv("data/county_voters_2016.csv")
county_voters_2018 <- read_csv("data/county_voters_2018.csv")
county_voters_2020 <- read_csv("data/county_voters_2020.csv")
county_voters_2022 <- read_csv("data/county_voters_2022.csv")
```

### Task 3: Using bind_rows

**Task** Combine the four files into a single data frame using the bind_rows function. Add a description of what this code does to your reference notebook.

```{r}
# bind_rows with list
county_voters_combined <- bind_rows(list(county_voters_2016, county_voters_2018, county_voters_2020, county_voters_2022))
View(county_voters_combined)
```

### Task 4: Asking a question

**Task** Run the following code to filter on Frederick County's registration data and look at the results. What stands out to you about the changes in party registration over time?
**Answer** 
Looking at the data below, There seems to be a decrease in voters in the Lib and Grn party in 2020 which at first I was thought was because of the pandemic. The interesting part is that the other parties did not decrease but rather increased despite the pandemic. Another thing I noticed is that the OTH category increased over time, but specifically between 2020 and 2022 in increased exponentially more than previous years. 
```{r}
county_voters_combined %>% filter(COUNTY == 'Frederick')
```

## Joining data

### Task 5: Loading population data

**Task** Load the Maryland 2020 county population data 

```{r}
maryland_population <- read_csv('data/maryland_population_2020.csv')
view(maryland_population)
```

### Task 6: Join population and voter data

**Task** Run the following code to join Maryland population data and the 2022 voter registration data together using the common column county as the key. How many rows are there now? How many *should* there be? 
**Answer**
There are 24 rows. There should be 23, because there are 23 counties in Maryland. There is probably a duplicate in the data that the function we just used was not able to combine or detect.

```{r}
# with nrow included to show row total
maryland_population %>% left_join(county_voters_2022, by="COUNTY") %>% nrow()
```

### Task 7: Looking at Joined Data

**Task** Examine the combined data

```{r}
# without nrow
maryland_population %>% left_join(county_voters_2022, by="COUNTY")
```

### Task 8: Assign Joined Data to New Dataframe

**Task** Create a new dataframe with population and voter data joined together. Add a description of what this code does to your reference notebook.

```{r}
maryland_population_with_voters <- maryland_population %>% left_join(county_voters_2022)
```

### Task 9: Add the Demographic Data

**Task** Load the Maryland demographic data and examine it. What kind of information does it contain?
**Answer** 
This contains information about demographics in the population of each Maryland county. It also contains geocodes for each county.
```{r}
maryland_demographics <- read_csv('data/maryland_demographics.csv')
view(maryland_demographics)
```

### Task 10: Join demographic data to combined voting/population data

**Task** Join the demographic data to the our combined voting/population data. To do this, we need to specify the join columns since the names aren't the same in the two dataframes.

```{r}
maryland_population_with_voters_and_demographics <- maryland_population_with_voters %>% left_join(maryland_demographics, by=c("COUNTY"="NAME"))
```

### Task 11: Asking Demographic Questions

**Task** Using mutate, let's find the county with the highest 18+ Black population as a percentage of all population 18 and over and also the percentage of Democrats in that county. Put your answers below. 
**Answer**
The county's with highest 18 plus Black population and Democrats is Somerset. Other counties with close percentages of Black population and Democrats is Prince George's and Charles county.

```{r}
maryland_population_with_voters_and_demographics %>%
  mutate(pct_black_18_plus = (pop_black/pop_18_over)*100, pct_dems = (DEM/Total)*100) %>% 
  arrange(desc(pct_black_18_plus)) %>% 
  select(COUNTY, pct_black_18_plus, pct_dems)
```

### Task 12: Asking Demographic Questions

**Task** Now let's find the county with the highest 18+ Asian population as a percentage of all population 18 and over and also the percentage of Democrats in that county. How do these results compare to the previous task? Put your answers below. 
**Answer**
The population of Asians is way less than the Black population and they are mostly concentrated in Howard and Montgomery county. This data shows that the Asian demographics, even if they were all democrats, would not make up a big amount of Democratic voters in each of these counties.
```{r}
maryland_population_with_voters_and_demographics %>%
  mutate(pct_asian_18_plus = (pop_asian/pop_18_over)*100, pct_dems = (DEM/Total)*100) %>% 
  arrange(desc(pct_asian_18_plus)) %>% 
  select(COUNTY, pct_asian_18_plus, pct_dems)
```

### Task 13: Fixing Join Problems

Sometimes joins look like they should work but don't. Often this is due to the two columns you're joining on having different data types: joining a <chr> column to a <dbl> column, for example. Let's walk through an example of that using some demographic data by zip code.

**Task** Run the following code to load the Zip Code Tabulation Area data for Maryland. What's the datatype of the ZCTA5N column? 
**Answer**
It is the <dbl> data type. We will need to change this to a <chr> in order to join it to other datasets.
```{r}
maryland_zcta <- read_csv('data/maryland_zcta.csv')
glimpse(maryland_zcta)
```

The column we're interested in, ZCTA5N, is a <dbl> column - it's a number. We want it to be a <chr> column - text.

**Task** Run the following code to change the datatype of ZCTA5N from numeric to character. What's the datatype of ZCTA5N? 
**Answer**
Now it is a <chr> column and now we can join it with the other datasets.

```{r}
maryland_zcta <- maryland_zcta %>% mutate(ZCTA5N = as.character(ZCTA5N))
```

Now we can join this dataframe to other zip code data where the zip code column is text, not numbers.
