---
title: "lab_03"
author: "Derek Willis"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv")
# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
```
```{r}
# cleaning function
cleaned_md_grants_loans <- md_grants_loans %>%
  clean_names()

# display the cleaned dataset
cleaned_md_grants_loans

cleaned_payments <- payments %>%
  clean_names()

  
cleaned_payments  
```



## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I found 54 records, including any variation of "pringe george county school" and of the education board.

```{r}
pg_schools <- cleaned_payments %>% 
  
filter(str_detect(vendor_name, "PRINCE GEORGES COUNTY SCHOOL") | str_detect(vendor_name, "BOARD OF EDUCATION OF PRINCE GEORGES COUNTY MD
") | str_detect(vendor_name, "PRINCE GEORGES PUBLIC SCHOOLS") | str_detect(vendor_name, "PRINCE GEORGES PUBLIC SCHOOL") | str_detect(vendor_name, "PRINCE GEORGES COUNTY PUBLIC SCHOOLS") | str_detect(vendor_name, "BOARD OF EDUCATION OF PRINCE GEORGE'S COUNTY MD
")) 




```
```{r}

```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: The fiscal year with the largest total amount is 2009 which had 8 total payment. Most of the years listed are not that recent. The most recent year in the data is 2020 and that year is near the bottom of total payments.

```{r}
pg_schools %>% 
  group_by(fiscal_year) %>% 
  summarize(count = sum(amount)) %>% 
  arrange(desc(count))
```
Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3: You would have to ensure that all the names of one vendor are the same all throughout the data. Here are three examples of names that could be combined.
1. 	PLANO-COUDON LLC and 	PLANO-COUDON, LLC
2. 

```{r}
cleaned_payments %>% 
  filter(str_detect(agency_name, "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND"))%>%
  
  group_by(agency_name, vendor_name, vendor_zip) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: I am running out of time but for this question I would search the use the 'filter' function to find what state grant money is going to 'stem cell research projects' I would then I would total up all the money for each fiscal year by grouping, summarazing and arranging. 

```{r}
md_grant_loans%>%
  mutate(description = str_to_upper(description)) %>%
  filter(str_detect(description,"STEM CELL RESEARCH GRANT"))%>%
  group_by(fiscal_year)%>%
  summarize(total = sum(amount))
```
